# Figure 3: Species composition

To illustrate how species composition changes with habitat types, we make a variety of plots:

1. The top-5 taxonomic families per habitat type.
1. The proportions of birds that belong to a selection of families that are representative for the entire Dutch landscape.

To quantify the necessary variables we rely mostly on the point-transect counts from Sovon as these cover all species (instead of the waterbird counts).

## Processing environment

```{r}
library(bioRad)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(ggpointdensity)
library(patchwork)
library(GGally)
library(tibble)
library(ggpubr)
library(forcats)
```

We load the study PPI and proportions of taxonomic groups in the counts.

```{r}
ppi <- readRDS("data/processed/composite-ppis/500m/201712312305.RDS")
wb_props <- readRDS("data/processed/sovon/wb_props.RDS")
ptt <- readRDS("data/processed/sovon/ptt.RDS")
ptt_props <- readRDS("data/processed/sovon/ptt_props.RDS")

ppi$data@data %>%
  # left_join(wb_props, by = c("wb_area_nr" = "area_nr")) %>%
  left_join(ptt_props, by = c("ptt_route" = "route")) %>%
  filter(coverage > 0,
       class != 1,
       total_biomass > 0,
       dist_radar < 66000,
       urban < 0.1) %>%
  identity() -> data
```

We select only those PPI pixels (and their corresponding counts) which are covered for >99% by a single habitat type.

```{r}
landuse_limit <- 0.99
```

## Characteristic families

Geese, ducks, pigeons, thrushes, tits, waders, crows and finches represent a substantial part of the Dutch avifauna across a large range of body sizes. We select these to illustrate how the taxonomic proportions change with habitats.

```{r}
selected_families <- c("Geese", "Ducks", "Pigeons", "Thrushes", "Tits", "Waders", "Crows", "Finches")

data %>%
  dplyr::select(all_of(c(colnames(ptt_props)[3:45], "semiopen", "forests", "wetlands", "waterbodies", "agricultural"))) %>%
  pivot_longer(cols = all_of(colnames(ptt_props)[3:45]), names_to = "family", values_to = "family_prop") %>%
  pivot_longer(cols = all_of(c("agricultural", "semiopen", "forests", "wetlands", "waterbodies")), names_to = "landuse", values_to = "landuse_prop") %>%
  drop_na() %>%
  filter(family %in% selected_families,
         landuse_prop >= landuse_limit) %>%
  group_by(family, landuse) %>%
  summarise(mean_family_prop = mean(family_prop), .groups = "drop_last") %>%
  pivot_wider(names_from = landuse, values_from = mean_family_prop) %>%
  mutate(family_sc = case_when(family == "Geese" ~ "Anatidae",
                               family == "Ducks" ~ "Anatinae",
                               family == "Pigeons" ~ "Columbidae", 
                               family == "Thrushes" ~ "Turdidae",
                               family == "Tits" ~ "Paridae",
                               family == "Waders" ~ "Charadrii",
                               family == "Crows" ~ "Corvidae",
                               family == "Finches" ~ "Fringillidae")) %>%
  identity() -> characteristic_families
characteristic_families
```

## Mean habitat biomass and weight

We calculate the mean biomass for all areas covered for >99% by a single habitat type and calculate mean bird weight for these areas from the point-transect counts.

```{r}
data %>%
  dplyr::select(all_of(c("total_biomass", "ptt_weighted_mean_weight", "semiopen", "forests", "wetlands", "waterbodies", "agricultural"))) %>%
  pivot_longer(cols = all_of(c("semiopen", "forests", "wetlands", "waterbodies", "agricultural")), names_to = "landuse", values_to = "proportion") %>%
  filter(proportion >= landuse_limit) %>%
  mutate(total_biomass = total_biomass / 1000) %>%
  group_by(landuse) %>%
  summarise(mean_biomass = mean(total_biomass),
            mean_weight = mean(ptt_weighted_mean_weight, na.rm = TRUE),
            .groups = "drop_last") %>%
  arrange(desc(mean_weight)) %>%
  identity() -> habitat_mass
habitat_mass
```

## Sorting of families by weight

Now we sort families according to weight

```{r}
ptt %>%
  distinct(species, .keep_all = TRUE) %>%
  group_by(familyvernacular) %>%
  summarise(mean_weight = mean(mean_weight), .groups = "drop_last") %>%
  filter(familyvernacular %in% selected_families) %>%
  arrange(desc(mean_weight)) %>%
  rowid_to_column()-> family_weights

characteristic_families %>%
  left_join(family_weights, by = c("family" = "familyvernacular")) %>%
  arrange(desc(mean_weight)) -> characteristic_families

characteristic_families$rank <- factor(characteristic_families$mean_weight, ordered = TRUE, levels = characteristic_families$mean_weight)
```

## Prepare plot

And plot a simple barplot figure.

```{r}
characteristic_families %>%
  pivot_longer(cols = !c("family", "rowid", "mean_weight", "rank", "family_sc"), names_to = "landuse", values_to = "family_prop") -> cf_long

colors <- c("urban" = "#94346E", "agricultural" = "#73AF48", "semiopen" = "#EDAD08", "forests" = "#0F8554", "wetlands" = "#38A6A5", "waterbodies" = "#1D6996",
            "dist_urban" = "#94346E", "total_biomass" = "#CC503E", "dist_radar" = "#666666")
landuse_names <- c("agricultural" = "Agricultural", "semiopen" = "Semi-open", "forests" = "Forests", "wetlands" = "Wetlands", "waterbodies" = "Water bodies")

plot_family_proportions <- function(props, which) {
  landuse_labeller <- function(variable, value) {
    label <- paste0(landuse_names[value], " ", round(habitat_mass[habitat_mass$landuse == which, ]$mean_weight, digits = 0), "g")
    return(label)
  }
  
  proportion_labeller <- function(value) paste0(value * 100, "%")
  
  props %>%
  filter(landuse == which) %>%
  ggplot() +
    geom_col(aes(x = family_prop, y = reorder(family, mean_weight), alpha = family_prop), fill = colors[which]) +
    labs(x = "Proportion of birds", y = "Taxonomic group") +
    coord_cartesian(expand = FALSE, xlim = c(0, ceiling(max(cf_long$family_prop) * 100) / 100)) +
    scale_x_continuous(breaks = c(0, 0.1, 0.2, 0.3), labels = proportion_labeller) +
    scale_alpha_continuous(range = c(0.4, 1)) +
    facet_wrap(vars(landuse), strip.position = "top") +
    theme_classic(base_size = 12) +
    theme(legend.position = "none",
          axis.title.y = element_blank(),
          strip.background = element_blank(),
          strip.text.x = element_blank())
}

landuses <- c("agricultural", "semiopen", "forests", "wetlands", "waterbodies")  # Same order as model outputs

plots <- mapply(function(landuses) { plot_family_proportions(cf_long, landuses) }, landuses = landuses,
            SIMPLIFY = FALSE)

# characteristic_families$title <- "Mean weight (g)"
# ggplot(characteristic_families) +
#   geom_col(aes(x = mean_weight, y = reorder(family, mean_weight))) +
#   labs(x = "Mean weight (g)", y = "Taxonomic group") +
#     coord_cartesian(expand = FALSE) +
#     scale_x_continuous(breaks = c(0, 1000, 2000, 3000)) +
#     facet_wrap(vars(title), strip.position = "top") +
#     theme_classic(base_size = 12) +
#     theme(legend.position = "none",
#           axis.title.y = element_blank(),
#           strip.background = element_blank(),
#           strip.text.x = element_text(hjust = 0, size = 12, family = "Helvetica", face = "bold")) -> plots[[6]]
p <- wrap_plots(plots, ncol = 1)
p
```
```{r}
p_models <- readRDS("data/plots/paper/fig_landscapes.RDS")

p_models[[1]] + p_models[[2]] + p_models[[3]] + p_models[[4]] + p_models[[5]] +
  plots[[1]] + plots[[2]] + plots[[3]] + plots[[4]] + plots[[5]] +
  plot_layout(widths = c(1, 1), nrow = 5, byrow = FALSE) & 
  theme(axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 12),
        axis.text.y = element_text(size = 10)) -> p_landscape_characterisation

ggsave(filename = "data/plots/paper/fig2_landscape.pdf", plot = p_landscape_characterisation, width = 14, height = 25, dpi = 300, units = "cm")
p_landscape_characterisation
```

```{r}
widths <- c(1, 1)
(p_models[[1]] / plots[[1]] * theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 10)) +
   plot_layout(widths = widths, ncol = 2)) /
  (p_models[[2]] / plots[[2]] * theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 10)) +
     plot_layout(widths = widths, ncol = 2)) /
  (p_models[[3]] / plots[[3]] * theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 10)) +
     plot_layout(widths = widths, ncol = 2)) /
  (p_models[[4]] / plots[[4]] * theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 10)) +
     plot_layout(widths = widths, ncol = 2)) / 
  (p_models[[5]] / plots[[5]] * theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 12), axis.text.y = element_text(size = 10)) +
     plot_layout(widths = widths, ncol = 2)) -> p_landscape_characterisation

p_landscape_characterisation
```

And we save it as .pdf for editing in Illustrator.

```{r}
ggsave(filename = "data/plots/paper/fig2_landscape.pdf", plot = p_landscape_characterisation, width = 14, height = 25, dpi = 300, units = "cm")
```
